---
title: "Fitness consequences of the selfish supergene _Segregation Distorter_"
author:
- Heidi W.S. Wong\textsuperscript{1} and Luke Holman*\textsuperscript{1}
- '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}'
bibliography: references.bib
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex_header.tex
  html_document: default
  word_document: default
fontsize: 12pt
csl: the-american-naturalist.csl
abstract: Segregation distorters are selfish genetic elements that subvert Mendelian inheritance, often by destroying gametes that do not carry the distorter. Simple theoretical models predict that distorter alleles will either spread to fixation, or stabilise at some high intermediate frequency. However, many distorter alleles defy these predictions, suggesting that we have yet to discover all salient evolutionary forces acting on distorter alleles. Here, we measured the fitness of \textit{Drosophila melanogaster} adults and juveniles carrying zero, one or two copies of three different variants of the naturally-occurring supergene \textit{Segregation Distorter} (\textit{SD}), in order to investigate why \textit{SD} remains relatively rare despite its strong distortion. First, we show that the three variants differ in the severity and dominance of the fitness costs they impose on carrier individuals. Second, we found instances in which \textit{SD} parents produce less fit offspring, suggesting that \textit{SD} alleles have non-genetic, transgenerational costs. Third, we found an \textit{SD} variant that altered the offspring sex ratio, perhaps due to off-target effects of \textit{SD} on the Y chromosome. Finally, we use an evolutionary model to investigate the effects of transgenerational costs and sex ratio effects on the evolutionary dynamics of \textit{SD}, and show that these previously undocumented costs potentially explain the puzzingly paucity of \textit{SD} alleles in wild \textit{D. melanogaster} populations. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(xtable)
options(xtable.comment = FALSE)

table1 <- read.csv("../figures/Table1.csv")
table1$Difference <- gsub(" ", "", table1$Difference)
table1$Difference <- gsub("to", " to ", table1$Difference)
table1$Difference <- gsub("\\(", " \\(", table1$Difference)
table1$Comparison <- gsub("-", "→", table1$Comparison)

table2 <- read.csv("../figures/Table2.csv")
table2$Difference <- gsub(" ", "", table2$Difference)
table2$Difference <- gsub("to", " to ", table2$Difference)
table2$Difference <- gsub("\\(", " \\(", table2$Difference)
table2$Comparison <- gsub("-", "→", table2$Comparison)


table1 <- xtable(table1, label="tb:expt1", 
                 size = "footnotesize", digits = c(0,0,0,0,0,1,3),
table2 <- xtable(table2, label="tb:expt2", 
                 size = "footnotesize", digits = c(0,0,0,0,1,3),
                 caption = "\\footnotesize{List of the all the notable differences between groups in Experiment 2 (posterior probability $<$0.05; see Table S10 for the remaining results). For each group, we list the sex of the focal larvae, their genotype (\\textit{SD} or \\textit{CyO}), and the parent that carried \\textit{SD} (mother or father). The difference in means is expressed in \\% larvae surviving; other details are as in Table 1.}") 
```

\maketitle\unmarkedfntext{\textsuperscript{1}School of BioSciences, The University of Melbourne, Victoria, Australia.}

\textbf{Keywords: gene drive, meiotic drive, population genetic model, selfish genes, \textit{t} paradox}.


\newpage
# Introduction
Segregation distorters are genetic elements that manipulate meiosis or gametogenesis to increase their representation in the gametes of heterozygotes above the usual 50% [@burt2006; @lindholm2016]. Because of this bias in transmission, segregation distorters are predicted to spread rapidly to fixation assuming that individuals carrying the distorter are equally fit as non-carriers [@bruck1957]. Even if the distorter reduces the fitness of individuals that carry it, it can still be favoured by selection provided that its individual-level fitness costs are outweighed by the within-individual advantage conferred by segregation distortion [@lindholm2016]. For this reason, there is currently great interest in developing natural or artificial segregation distorters that could propagate human-beneficial alleles through a wild animal or plant population, for example to introduce malaria resistance alleles into wild mosquito populations [@gantz2015]. In addition to their promise in applied science, the study of segregation distorters has led to multiple advances in our understanding of evolution, genetics, and speciation [@rice2013; @lindholm2016; @manser2017; @lin2018; @verspoor2018].

A long-standing puzzle, sometimes called the '$t$ paradox' after the well-studied mouse segregation distorter $t$ [@carroll2007], concerns the failure of many segregation distorters to invade populations as well as simple models predict that they should [@bruck1957; @lewontin1968; @charlesworth1978; @taylor2002; @holman2015]. For example, Bruck [-@bruck1957] showed that a homozygous lethal distorter that is transmitted to a faction $k$ (0.5 < $k$ $\ge$ 1) of the offspring of heterozygous males should reach an equilibrium allele frequency of $\frac{1}{2} - \sqrt{(k(1-k))} / 2k$, if one makes no further complicating assumptions. This prediction is 38.5% if $k$ = 0.95, which is substantially higher than the allele frequencies observed for real-world homozygous lethal segregation distorters with $k \approx$ 0.95, such as the mouse $t$ haplotype [5-14%; @ardlie1998] and the _Segregation Distorter_ gene complex of _Drosophila melanogaster_ fruitflies [0-8%; @brand2015]. Several later models sought to resolve this discrepancy by adding additional assumptions to the basic population genetic model. For example, Lewontin [-@lewontin1962] argued that population structure can reduce the equilibrium frequency of a distorter allele [see also @bull2019], and Lewontin [-@lewontin1968] showed that strong, partially recessive fitness costs can maintain a distorter allele at low, stable frequencies. Additionally, males carrying segregation distorters often perform worse in sperm competition due to the loss of half their gametes, and there is some evidence that this can create frequency-dependent selection against distorters [@taylor2002; @holman2015]. 

Here, we measure the magnitude and dominance of the fitness costs to larvae and adults carrying either one or two copies of _Segregation Distorter_ (_SD_) in _D. melanogaster_. _SD_ is a gene complex or 'supergene' [@thompson2014] composed of several linked loci on chromosome 2 (an autosome) that causes strong segregation distortion in heterozygous males by disrupting the development of non-SD-carrying spermatids [reviewed in @larracuente2012]. The _SD_ supergene contains an 'insensitive' allele at the _Responder_ locus (_Rsp_), while most chromosomes that lack _SD_ carry a 'sensitive' _Rsp_ allele that makes them susceptible to distortion. Chromosomal inversions in the _SD_ region help to keep the component loci in linkage, preventing the creation of recombinant 'suicide chromosomes' in which the insensitive _Rsp_ allele linked to _SD_ is replaced by a sensitive allele. The threat of suicide chromosomes selects for reduced recombination, and indeed _SD_ is often surrounded by a large non-recombining region [\textit{c}. 10% of the genome; @presgraves2009] containing deleterious mutations [@temin1984; @larracuente2012; @brand2015]. _SD_ is thought to have a single evolutionary origin around 38,000 years ago [@brand2015], though it has since diversified into multiple variants which differ in their inversions and associated mutation load [@larracuente2012; @presgraves2009; @brand2015]. In some populations, _SD_ chromosomes are present at low, stable frequencies that suggest balancing selection [e.g. 0-8% in 14 populations; @brand2015], although high and unstable allele frequencies have also been reported: one _SD_ variant increased in frequency from 17% to 98% over 23 years in Wisconsin [-@temin1984].

The evolutionary dynamics of segregation distorters depend strongly on the fitness of drive-carrying individuals. Negative frequency-dependent selection is of particular interest because it can maintain balanced polmorphisms in the face of strong distortion [@holman2015]. If the fitness of carriers declines as the distorter allele becomes more common, the costs and benefits of the distorter can balance, preserving both distorting and non-distorting alleles. Recessive fitness costs create such selection, because distorter homozygotes become more common as the distorter invades. However, some distorters appear healthy when homozygous [@temin1984; @price2012], meaning that recessive costs cannot provide a complete answer to the $t$ paradox. Additionally, models [e.g. @bruck1957; @lewontin1968] demonstrate that homozygote lethality is insufficient to explain the low allele frequencies of strong distorters like _SD_ or $t$.

Here, we focus on the three best-studied _SD_ variants: SD-5, SD-72, and _SD-Mad_ [all originally from Wisconsin; @sandler1959]. _SD-5_ carries a different set of inversions from the other two and is reportedly homozygous lethal [@larracuente2012], while some SD-72- and SD-Mad-type alleles are reported to be fit as homozygotes [@temin1984]. To our knowledge, the relative fitness of _SD_ heterozygotes has not been measured, though this parameter is crucial to the evolutionary dynamics of a distorter allele [@lewontin1968]. As well as measuring these costs, we wished to verify older reports [@hiraizumi1967; @denell1969] that males carrying _SD_ produce offspring with an atypical sex ratio, and to investigate theoretically how such a sex ratio bias would affect the evolution of _SD_. Lastly, we tested for non-genetic fitness effects of _SD_, e.g. mediated by parental effects or genomic imprinting, and use models to investigate the effects on _SD_ evolution. We discuss the evolution of _SD_ in wild populations in light of our empirical and theoretical findings.

# Methods

## Fly stocks
All flies were reared at 25^o^C under natural light (_c._ 14h day length) in 25mm plastic vials containing food medium (yeast-soy-cornmeal-agar-corn syrup). All stocks were obtained from the Bloomington Drosophila Stock Centre unless otherwise stated (SD stock numbers: 64322, 64324, and 64323). 

In order to generate a non-SD reference allele which also allowed us to visually distinguish flies carrying 0, 1 or 2 copies of _SD_, we created a stock carrying an isogenic copy of chromosome 2 that was labelled with one recessive and one dominant phenotypic marker. The recessive marker was a mutant allele of _bw_ encoding brown eye colour (obtained from a teaching laboratory in Melbourne; unknown origin), while the dominant marker was the transgene _Ubi-GFP_ (stock 5826), which expresses green fluorescent protein (GFP) throughout the body. To recombine these markers, we crossed F1 _bw/Ubi-GFP_ females to _bw_ males and collected male progeny expressing brown eyes and GFP. From these recombinants, we selected a single male and crossed him to a female carrying wild-type X chromosomes (one from the _bw_ stock and one from the _SD-72_ stock) as well as the balancer chromosome _SM5_, collected +/+; _bw-GFP/SM5_ progeny, and crossed them to create what we hereafter call the _bw-GFP_ stock.

In the adult fitness assays, we used opposite-sex _bw_ individuals as mates, and _Gla/CyO_ individuals (stock 44227) as same-sex competitors. The offspring of _Gla/CyO_ flies express a dominant mutant phenotype distinguishing them from the offspring of the focal flies.

Lastly, the three SD-bearing Bloomington stocks had different balancer chromosomes (SD-5 used _CyO_, _SD-72_ used _SM5_, and _SD-Mad_ was not balanced), so we first standardised the balancer to _CyO_ (from the _Gla/CyO_ stock) to remove this potential confounding effect. We then crossed _SD/CyO_ progeny to the _bw-GFP_ stock to create _SD/bw-GFP_ individuals. 

## Experiment 1

### Experimental crosses
We performed four types of experimental crosses for each of the three _SD_ chromosomes (Figure 1). In Cross 1, we mated two _SD/bw-GFP_ flies, yielding offspring carrying 0, 1 or 2 _SD_ chromosomes. In Cross 2, we mated _SD/bw-GFP_ females to _bw_ males, yielding offspring carrying 0 or 1 _SD_ chromosomes. Cross 3 was the reciprocal: a _bw_ mother and _SD/bw-GFP_ father. Lastly, to measure the baseline fitness of the _bw-GFP_ chromosome, we mated two _bw-GFP_ flies (Cross 4). 

All of these crosses were performed in parallel on a common cohort of flies under identical conditions in a randomised order, minimising confounding effects. We ran all four crosses (and their associated fitness assays; see below) in each of three experimental blocks, with equal representation of crosses within blocks. We measured three components of fitness: survival rate from first-instar larva (hereafter 'L1 larvae') to adult, adult male competitive fertilisation success, and adult female fecundity following social interaction. For brevity, we term these juvenile, male, and female fitness. We also recorded the sex ratio of individuals that reached adulthood in the juvenile fitness assay.

### Juvenile fitness and sex ratio assays
Mated females from the four experimental crosses were placed separately onto egg collection plates (grape-agar medium with live yeast) for 24h, then removed. We waited 24h, then collected L1 larvae and sorted them by GFP phenotype. The reason for beginning the assay with L1 larvae, not eggs, was that we could correctly classify the GFP phenotype of L1 larvae (100/100 successes in a pilot) but not eggs. We placed the sorted larvae in fresh vials in groups of up to 100. It was difficult to obtain 100 larvae for every class of progeny because some progeny classes are rare due to segregation distortion and/or mortality in the embryonic stage. We subsequently quantified juvenile fitness and the sex ratio by counting, sexing, and phenotyping the adults that eclosed from these vials. 

### Adult female and male fitness assays
Flies that survived to adulthood in the juvenile fitness assay were sorted by phenotype/genotype into single-sex vials, left to mature for 48-72h, and then used in adult fitness assays. 

To measure female fitness, we placed 5 same-genotype females in an 'interaction vial' with 15 _bw_ males and 10 _Gla/CyO_ females (all flies were 48- to 72-hour-old virgins), and allowed them to interact for 48h to facilitate mating, courtship, behavioural interactions such as harassment, and competition for food. We then recorded the number of surviving focal females, and moved them to a new yeasted food vial (without the non-focal flies), where they oviposited for 24h. We then removed the females and counted the number of larvae eclosing from their eggs, and used this as our measure of female fitness. Thus, our measure of female fitness measure is the product of female fecundity, the proportion of eggs that are fertilised, and survival rate in the zygote-to-L1 stage. 

To measure male fitness, we placed 5 same-genotype males in an interaction vial with 15 _bw_ females and 10 _Gla/CyO_ males (again, all flies were 48- to 72-hour-old virgins), where they interacted and mated for 48h. We then moved all surviving individuals (focal and non-focal) to a new food vial where they continued to interact and oviposit for 24h. We then removed all adults and allowed their offspring to develop to adulthood, then counted the number of progeny sired by the focal males and the competitor _Gla/CyO_ males. We used the proportion of progeny sired by the focal males as a measure of adult male fitness. This fitness measure encompasses pre- and post-copulatory sexual selection, as well as the survival rate of focal males’ offspring relative to those of _Gla/CyO_ males. 

### Limitations of Experiment 1’s juvenile fitness assay
Upon phenotyping adult flies emerging from Crosses 1-4, we observed unexpected recombination between the _bw_ and _Ubi-GFP_ loci for the _SD-72_ and _SD-Mad_ (but not SD-5) chromosomes [we had assumed that _SD_ chromomes would be largely non-recombining; presgraves2009]. Specifically, in Cross 2, some GFP-negative larvae developed brown eyes, and some GFP-positive ones developed red eyes, indicating recombination in the _SD/bw-GFP_ mother (recombinants were never seen in Cross 3, because there is no recombination in male _Drosophila_; this lets us rule out GFP detection errors). The proportion of recombinant adults in Cross 2 was 3.6% (95% CIs: 2.4-4.9%) for SD-5, 36.1% (33-39%) for SD-72, and 32.8% (30-36%) for SD-Mad. The _bw_ locus is at the terminal end of the right arm of chromosome 2 (2R), and _SD-5_ is distinguished from the other two variants by an additional inversion on 2R; we therefore hypothesise that the _Ubi-GFP_ transgenic insertion lies somewhere on 2R between the _SD_ complex and _bw_, probably close to the SD-5-specific inversion [Figure 1 in @larracuente2012]. As a consequence of this unexpected recombination we cannot be certain how many larvae of each genotype were present at the start of the juvenile fitness assay for Cross 2, at least for _SD-72_ and _SD-Mad_ -- we simply removed the recombinant individuals from the dataset, and made the simplistic assumption that all of the larvae that did not reach adulthood were non-recombinants. We interpret the relevant part of the Results in light of the resulting bias. This limitation is offset by data from Experiment 2 (which does not rely on these markers, and uses a non-recombining balancer chromosome), as well as data from Cross 3 (since there is no recombination in male _Drosophila_).

Additionally, for Cross 1, individuals carrying 0 or 1 _SD_ chromosomes were phenotypically indistinguishable until they reached adulthood and developed eyes, and so we simply the measured the survival rate of a mixed pool of larvae carrying either 0 or 1 _SD_ alleles. Most larvae in this pool carried 1 _SD_ allele, rather than 0, because of segregation distortion. Specifically, the proportion of _SD_ progeny in the pool will be $1 / (k + 2(1-k))$, or 95.2% for $k$ = 0.95. This limitation is offset by data from Crosses 2 and 3 and Experiment 2.

## Experiment 2
Experiment 2 was designed to measure the direct and transgenerational effects of _SD_ on sex-specific larval survival, and to address the limitations of Experiment 1. Experiment 2 used the transgenic construct _P{Sxl-Pe-EGFP.G}G78b_ (extracted from stock 24105, backcrossed into the $w^{1118}$ genotype for 5 generations, and made homozygous), which allows discrimination of males and females at the egg stage [female-destined embryos express GFP while males do not; @thompson2004]. We conducted six types of crosses using parents bred at standardised density: in each cross, one parent was _SD/CyO_ and the other was homozygous for _P{Sxl-Pe-EGFP.G}G78b_; we performed this cross with the three _SD_ variants, with either the mother or the father providing _SD_ (10-24 replicates per cross). We then collected embryos of both sexes (mean: 48 embryos per sex per cross), placed them in single-sex vials to develop, and then counted and phenotyped the eclosing adults to infer the survival rates of different progeny classes. 

## Statistical analysis
We analysed Experiment 1 using Bayesian hierarchical models implemented in the R package `brms` [@burkner2017]. The data on juvenile fitness, male fitness, and adult sex ratio were treated as binomially distributed, and we fit 'vial' as a random effect to account for nonindependence among individuals from the same vial. Female fitness was modelled using the negative binomial distribution, since the data were overdispersed counts. We verified model fit using posterior predictive checks [@gelman2006], as shown in the Online Supplementary Material. 

For hypothesis testing, we calculated the posterior differences between group means for pairs of means (contrasts) that are informative for this study (e.g. the difference in fitness between individuals with 0 or 1 _SD_ allele, or individuals that received _SD_ from their father versus their mother). Differences for which most of the posterior lies far from zero are highlighted in Tables 1 and 2. We also calculated the posterior probability that the group with the larger posterior mean actually has a smaller mean than the other group; this provides a metric with a similar interpretation to the more familiar $p$-value.

The aim of Experiment 2 is to estimate the proportion of _SD_ and non-SD male and female larvae that survive to adulthood. However, because the genotype of larvae could not be visually determined at the start of Experiment 2, we had to estimate the initial numbers of larvae belonging to each genotype in order to calculate a survival rate. For example, if we placed 50 larvae in a vial and 20 non-SD and 20 _SD_ individuals reached adulthood, we inferred the genotypes of the 10 dead ones in order to estimate the relative survival rates of _SD_ and non-SD individuals. This unmeasured variable depends on the gametes produced by the _SD/CyO_ parent. Because _SD_ is well-documented to only cause distortion in males [@larracuente2012], we assumed that the _SD/CyO_ mothers transmitted _SD_ to 50% of their progeny. We also assumed 50% transmission in _SD/CyO_ fathers (i.e. $k$ = 0.5), in light of evidence that _CyO_ carries an insensitive allele of _Rsp_ that makes it immune to segregation distortion [@ganetzky1977]. We then used a binomial random number generator with $p$ = 0.5 to 'guess' the genotypes of the dead larvae. Our sample size was sufficiently large that generating a new set of random numbers and re-running the model gave near-identical results, thanks to the law of large numbers. We also re-ran the model under the assumption that there is some segregation distortion in fathers [i.e. \textit{k} > 0.5, contradicting the evidence in @ganetzky1977], and found that all the key results did not change (see Online Supplementary Material).

## Population genetic model
Our experiments suggested that some _SD_ variants have parent-of-origin-specific effects on fitness and/or cause SD-carrying males to produce a biased offspring sex ratio. We therefore constructed a simple one-locus, two-allele population genetic model to examine the effect of these two factors on the evolution of _SD_. 

The model considers the spread of an autosomal segregation distorter in an infinitely large, panmictic population with discrete generations. We assume that individuals carrying two wild type alleles have a relative fitness of 1, while other genotypes potentially have relative fitness between 0 and 1. We kept track of the parental origin of the _SD_ allele in heterozygotes, to allow heterozygotes with a maternally-inherited _SD_ to have a different fitness than heterozygotes with a paternally-inherited _SD_, and thereby allow for the possibility that _SD_ has a parent-of-origin-specific effect on fitness. We assumed that male heterozygotes transmit the _SD_ allele to a fraction $(1 + K)/2$ of their offspring (where 0 < $K$ < 1), and produce a fraction $(1 + s)/2$ female offspring (0 < $s$ < 1), while all other genotypes were assumed to show normal Mendelian inheritance and a 50:50 offspring sex ratio. For example, a mating between a wild-type female and a male _SD_ heterozygote produces $(1 + K)(1 - s)/4$ heterozygote sons, $(1 + K)(1 + s)/4$ heterozygote daughters, $(1 - K)(1 - s)/4$ wild-type sons, and $(1 - K)(1 + s)/4$ wild-type daughters. Note that for convenience and consistency with other models, the model uses capital $K$ (range: 0-1, where 0 indicates no distortion and 1 complete distortion), rather than the lowercase $k$ discussed earlier (where 0.5 indicates no distortion and 1 complete distortion).

For each parameter space, we determined the evolutionary fate of an _SD_ allele in a starting population with 1% _SD_ alleles at Hardy-Weinberg genotype frequencies. We calculated the equilibrium genotype frequency using numerical simulations, since the analytical solution would be unwieldy. In each generation, we first multiplied the frequency of each genotype by its relative fitness (representing the combined action of natural and sexual selection across all life stages) and then renormalised the genotype frequencies to sum to one. We then determined the frequency of each of the possible mating types as the product of each possible pair of maternal and paternal genotype frequencies. From these, we determined the offspring genotype frequencies, and replaced the parental generation with the offspring. The simulation ran for 10,000 generations, or until the _SD_ allele went extinct (defined as reaching 0.001% frequency) or fixed (>99%).

## Availability of data and code
All our raw data plus the R code used for our analyses is provided as supplementary material, and can also be viewed online at [https://lukeholman.github.io/fitnessCostSD/](https://lukeholman.github.io/fitnessCostSD/).

# Results

## Experiment 1
Posterior estimates of mean fitness for each group are plotted in Figure 1. Table 1 lists notable pairwise differences between groups, Tables S1-S4 give sample sizes and simple summary statistics, and Tables S6-S9 give results for all the contrasts we examined.

\begin{figure}
\centering
\includegraphics{../figures/fig1.pdf}
\caption{\footnotesize{Posterior estimates of the group means for the four different response variables in Experiment 1, for each type of cross (x-axis), \textit{SD} variant (panels), and genotype (colours). Juvenile fitness was measured as % L1 larva-to-adult survival, female fitness as the estimated number of progeny produced per female, and male fitness as the siring success relative to competitior males. The thicker inner bar shows the region containing 50\% of the posterior, while the outer bar covers 95\% of the posterior. Tables S6-S9 give the accompanying statistcal results. Points labelled as carrying ``0 or 1'' \textit{SD} allele refer to cases where the genotype of the offspring could not be ascertained due to the brown eye marker not being expressed in larvae; most of these individuals probably carried 1 \textit{SD} allele because of segregation distortion.}}
\end{figure}

```{r results='asis'}
print(table1, booktabs=TRUE, size='footnotesize', include.rownames = FALSE, caption.placement = "top")
```

### Juvenile fitness
When collecting larvae we observed 40 L1 larvae homozygous for SD-5, and over 600 carrying two copies of SD-72, but none of these survived to adulthood. The smaller number for _SD-5_ indicates that most _SD-5_ homozygotes died before hatching, while _SD-72_ homozygotes tended to die between hatching and adulthood. By contrast, many larvae homozygous for _SD-Mad_ reached adulthood, and there was no detectable fitness effect of _SD-Mad_ on juvenile fitness even in homozygotes.

The limitations of this assay (see Methods) mean that Figure 1 might underestimate the survival rate of individuals carrying a maternally-inherited _SD_ allele, for _SD-72_ and _SD-Mad_. Therefore, we cannot be certain that there is no difference in juvenile fitness between individuals with an _SD_ mother versus an _SD_ father for _SD-72_ and _SD-Mad_.


### Sex ratio among individuals reaching adulthood
For crosses in which the father carried _SD-Mad_, the sex ratio of the emerging adults was significantly more female-biased than for crosses in which the mother carried _SD-Mad_, irrespective of offspring genotype. The results did not replicate earlier findings that the non-_SD_ offspring of _SD_ heterozygote fathers show a female-biased sex ratio [@hiraizumi1967; @denell1969]; indeed, there was a nonsignificant trend in the opposite direction for _SD-5_ (the posterior median % male offspring was 54% among the non-_SD_ offspring and 48% male among the _SD_ offspring; Figure 1).   

### Adult female fitness
Females homozygous for _SD-Mad_ had low fitness. Additionally, females were fitter if their mother, as opposed to their father, carried _SD-5_ or _SD-Mad_, irrespective of whether the offspring inherited _SD_.  

### Adult male fitness
Males homozygous for _SD-Mad_ had low fitness. We again observed evidence for non-genetic effects: for _SD-5_, males with a paternally-inherited _SD_ chromosome were substantially less fit than males with a maternally-inherited _SD_ chromosome. There was also a strong fitness cost of inheriting an _SD-5_ allele from either parent. Interestingly, males with an _SD-Mad_ father were fitter if they inherited _SD-Mad_ rather than the non-_SD_ allele.

## Experiment 2
Experiment 2 also suggested that _SD_ chromosomes can have both direct and transgenerational effects on L1 larva-to-adult survival (Figure 2; Table 2; full results in Tables S5 and S10). Male larvae with an _SD-5/CyO_ mother were significantly less likely to survive than those an _SD-5/CyO_ father, irrespective of whether the larva actually inherited _SD-5_ (the same was not true for the other _SD_ variants, suggesting that _SD-5_ not _CyO_ mediates this effect). Additionally, female larvae who inherited _SD-5_ from the mother survived less well than did females who inherited _CyO_, suggesting that a single copy of _SD-5_ reduces survival more than a single copy of _CyO_. The same effect was not observed for males, or for crosses in which _SD-5_ came from the father, possibly indicating that _SD-5_ has a sex- or parent-of-origin-specific effect on survival (we lack the sample size to be certain). By contrast, male larvae whose father carried _SD-Mad_ were fitter if they inherited _SD-Mad_ rather than _CyO_, suggesting that _SD-Mad_ is less harmful than _CyO_. Lastly, we observed some significant sex differences in survival for all three _SD_ chromosomes, with female larvae surviving better than male larvae. We did not find any evidence that the direct genetic effect of _SD_ on larval survival is sex-specific: the difference in survival rate for _SD_ and _CyO_ individuals was similar for males and females (Figure 2). 

\begin{figure}
\centering
\includegraphics{../figures/fig2.pdf}
\caption{\footnotesize{Posterior estimates of \% L1 larva-to-adult survival in Experiment 2 for each combination of offspring sex and genotype (x-axis), \textit{SD} variant (panels), and cross (colours). The thicker inner bar shows the region containing 50\% of the posterior, while the outer bar covers 95\% of the posterior. See Tables 2 and S10 for associated hypothesis tests. The model underlying this plot assumed fair meiosis ($k$ = 0.5) in \textit{SD/CyO} males; see Figure S1 for equivalent plots made using different assumed values of $k$.}}
\end{figure}

```{r results='asis'}
print(table2, booktabs=TRUE, size='footnotesize', include.rownames = FALSE, caption.placement = "top")
```


## Population genetic model
We first assumed that the _SD_ allele had no direct or transgenerational fitness costs (top left, Figure 3), which resulted in the invasion of _SD_ even if segregation distortion ($K$) was very weak. However, if the _SD_ allele caused males carrying it to produce a highly biased sex ratio (unrealistically high, based on our data), _SD_ required a higher $K$ to invade. The reason that a sex ratio bias for SD-carrying males hinders the spread of _SD_ is that autosomal loci usually maximise their fitness by producing a 50:50 sex ratio, due to 'Fisherian' selection on the sex ratio, which disfavours alleles causing unequal production of sons and daughters [@fisher1930]. In cases where the _SD_ allele was able to invade, it generally went to fixation: a balanced polymorphism of _SD_- and non-_SD_ was seldom observed. There was a small zone of polmorphism when drive was very weak and the sex ratio bias was very strong, due to the frequency dependence of the costs associated with the biased sex ratio (i.e. sex ratio-biasing alleles are more costly in populations where they are common, due to the increasing bias in the population sex ratio).

Secondly, when we assumed dominant costs such that all individuals with at least one _SD_ allele had a relative fitness of 0.8 (top second left, Figure 3), the _SD_ allele could still invade, though it needed a substantially higher transmission bias $K$. When _SD_ did invade, it again proceeded to fixation, except under unrealistically weak drive and extreme sex ratio bias. Notably, invasion was more difficult when _SD_ heterozygote males produce a female-biased rather than male-biased sex ratio. Such a female bias is detrimental to _SD_ because segregation distortion only happens in males, increasing the critical value of $K$ required for invasion. _SD_ invaded slightly more easily when _SD_ heterozygote males produced >50% sons, but not so easily as when there was no sex ratio bias (due to the fitness gains of extra transmission bias through sons being opposed by Fisherian sex ratio selection).

Thirdly, when we assumed that _SD_ is recessive-lethal but cost-free in heterozygotes (top second right, Figure 3), the _SD_ allele stabilised at high, intermediate frequencies for realistic (i.e. high) values of $K$ [as expected; @bruck1957]. This is because recessive fitness costs create negative frequency-dependent selection, halting the spread of the _SD_ allele once the costs experienced in homozygotes cancel out the effect of segregation distortion [@holman2015]. A female-biased sex ratio reduced the equilibrium frequency of _SD_ while a male-biased sex ratio had little effect, due to the opposing effects of Fisherian selection and the benefits of producing more sons (i.e. the sex in which distortion occurs).

Fourthly, we modelled a recessive-lethal _SD_ that reduces the relative fitness of heterozygotes to 0.8 (top right, Figure 3 - this assumption is probably the most realistic so far, based on our empirical findings). Here, the _SD_ allele only invaded when $K$ was high, and it stabilised at medium-high frequencies. Interestingly, _SD_ alleles that induced a male-biased sex ratio invaded for substantially lower $K$ and reached a higher equilibrium frequency for any given $K$ than those that did not affect the sex ratio. Presumably this occurred because when _SD_ is kept rare by its direct fitness costs, the population sex ratio stays close to 50:50, and so Fisherian sex ratio selection against _SD_ remains weak (while the benefits of extra transmission bias stay the same).

For all four of these scenarios, we produced similar graphs under the additional assumption that offspring suffer an additional cost when the _SD_ allele is inherited from a particular parent. In the middle row of Figure 3, genotypes carrying a paternally-inherited _SD_ allele have their fitness reduced by an additional 0.2, while in the bottom row, the same applies to genotypes with a maternally-inherited _SD_. Comparison of the three rows shows that these trans-generational costs further hamper the spread of _SD_, and that paternal costs are worse than maternal costs, because the _SD_ allele is inherited from fathers more often than mothers due to its male-limited distortion. By combining recessive lethality with some mixture of heterozygote fitness costs, sex ratio bias, or transgenerational costs, we could get _SD_ chromosomes to persist at low, stable frequencies as they often do in nature (e.g. the middle right panel of Figure 3 near $K$ = 0.95, which approximates the costs and $K$ value for _SD-5_). 


\begin{figure}
\centering
\includegraphics{../figures/fig3.pdf}
\caption{\footnotesize{The equilibrium frequency reached by the \textit{SD} allele, $\hat{p}(SD)$, depends on the strength of segregataion distortion ($K$ = 0 indicates fair meiosis, $K$ = 1 that heterozygotes transmit only the \textit{SD} allele) and the direction and strength of sex ratio bias in the progeny of \textit{SD} heterozygote males. The four columns make different assumptions about the fitness costs to individuals carrying the \textit{SD} allele (see Results), while the three rows assume either that \textit{SD} has no parent-of-origin-specific effects on fitness (top row), or that \textit{SD} is especially costly when paternally inherited (middle row) or maternally inherited (bottom row).}}
\end{figure}



# Discussion
Our results reaffirm that _SD-5_ and _SD-72_ are homozygous lethal. Most _SD-5_ homozygotes died while still in the egg, while _SD-72_ homozygotes died after hatching but before adulthood. Although cultures fixed for _SD-Mad_ can survive in the lab, we found that male and female _SD-Mad_ homozygotes had much lower fitness than the _bw-GFP_ competitor flies, suggesting that _SD-Mad_ would be effectively homozygous lethal in a natural population. The fitness costs to male adults were dominant for _SD-5_ but recessive for _SD-72_ and _SD-Mad_. By contrast, the fitness costs to female adults were recessive for all three _SD_ variants, illustrating that the dominance of the cost differs between sexes as well as between _SD_ variants. Although we did not observe any _SD_ variants that were fit as homozygotes in this study, it is possible that such variants exist; an _SD_ variant with inversions characteristic of _SD-72_ or _SD-Mad_ was observed to reach 98% frequency in a population in Wisconsin [@temin1984]. 

Interestingly, we found some evidence for costly non-genetic transgenerational effects associated with all three _SD_ variants, mediated either by parental effects [sensu @badyaev2009] or genomic imprinting [sensu @holman2014]. Firstly, female fitness was reduced among the non-SD offspring of _SD-5_ or _SD-Mad_ heterozygote fathers. One possible mechanism is that chromosomes that escape segregation distortion are epigenetically modified in ways that affect adult fitness; this mechanism is plausible because _SD_ seems to function by altering the chromatin of sensitive chromosomes [@larracuente2012]. Secondly, _SD-5_ was especially harmful to adult male fitness when paternally inherited, hinting at either genomic imprinting or a genotype-dependent paternal effect of _SD-5_ (either phenomenon would be consistent with the data). Thirdly, in Experiment 2, we found that male larvae whose mother carried _SD-5_ were less likely to reach adulthood than were male larvae whose father carried _SD-5_, irrespective of whether the larva inherited _SD-5_. This result again suggests that _SD-5_ has a transgenerational effect on offspring fitness, though puzzlingly the harmful effect was associated with mothers rather than fathers this time (possibly because Experiments 1 and 2 used a different non-_SD_ reference chromosome and genetic background). To our knowledge, all previous theoretical models of segregation distorters implicitly assume that transgenerational effects are absent, so we built a simple model incorporating the assumption that _SD_ alleles can have parent-of-origin-specific effects on fitness. The model showed that non-genetic transgenerational costs of _SD_ can reduce the invasion probability and equilibrium frequency of _SD_ [this is an example of evolution via indirect genetic effects, @wolf1998]. Thus, if segregation distorters commonly have transgenerational costs, such costs may help to explain the puzzling rarity of _SD_ [@brand2015] and other autosomal distorters such as the $t$-haplotype [@carroll2007] in spite of biased segregation.

We also observed that fathers heterozygous for _SD-Mad_ produced an excess of daughters, while _SD-5_ and _SD-72_ parents produced a similar sex ratio to controls. Our results thus differ from earlier studies of _SD-5_ and _SD-72_, which found an excess of daughters but only among the non-_SD_ progeny [@hiraizumi1967; @denell1969]. Larracuente and Presgraves [-@larracuente2012] proposed that Y-bearing spermatids might be eliminated in _SD_ males as a result of ‘collateral damage’ arising because of sequence homology between Y-linked loci and _Responder_, which could explain the shortage of sons that we observed. Additionally, we speculate that _SD_ might cause a parental effect that affects the relative survival rates of male and female progeny, for example by inducing epigenetic modifications that are more harmful to males than females. Our modelling results suggest that _SD_ alleles invade less easily, and reach a lower equilibrium frequency, when they cause male heterozygotes to produce a female-biased sex ratio. There are two reasons for this result: firstly, autosomal alleles that skew the sex ratio away from 50:50 are usually disfavoured by selection [@fisher1930], and secondly, _SD_ alleles can only distort segregation in sons. The model also showed that producing a male-biased sex ratio was disadvantageous for _SD_ alleles, except in populations where _SD_ was kept rare by its fitness costs. When _SD_ is rare, the population-wide sex ratio remains close to 50:50, reducing the Fisherian cost to _SD_ of producing extra sons. Assuming that other autosomal segregation distorters also cause imbalanced sex ratios, this finding may help to resolve the $t$-paradox in other systems.

Future studies could compete _SD_ alleles with differing costs, and differing cost dominance, in population cages. We predict that _SD_ alleles that have dominant costs will either fail to spread (if the costs are sufficiently high relative to the strength of segregation distortion, $k$), or will sweep to fixation, while alleles with recessive costs will potentially reach an evolutionary equilibrium. Similarly, we predict that the stability and allele frequencies of _SD_ chromosomes in natural populations will relate to their fitness costs in homozygotes and heterozygotes. _SD-5_ is more costly, has more dominant costs, and was rarer than other the other two variants in the original Wisonsin population [@temin1984], and it would be interesting to see if the frequencies of competing _SD_ variants can be similarly explained in other populations. Our results also have implications for the design of artificial gene drives [@lindholm2016]. For example, we suggest that it is worthwhile to measure the fitness of drive-carrying individuals' offspring (not just the fitness of the carriers themselves) when testing a newly-designed gene drive in the lab, since our model shows thattransgenerational costs can strongly influence the invasion success of a gene drive. 


# Acknowledgements

# References
<div id="refs"></div>

\newpage
# Online Supplementary Material
