---
title: "Fitness consequences of the selfish supergene _Segregation Distorter_"
bibliography: references.bib
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex_header.tex
fontsize: 12pt
csl: the-american-naturalist.csl
abstract: \noindent{Segregation distorters are selfish genetic elements that subvert Mendelian inheritance, often by destroying gametes that do not carry the distorter. Simple theoretical models predict that distorter alleles will either spread to fixation, or stabilise at some high intermediate frequency. However, many distorter alleles are less common than expected, suggesting that key evolutionary forces remain to be discovered. Here, we measured the fitness of \textit{Drosophila melanogaster} adults and juveniles carrying zero, one, or two copies of three different variants of the naturally-occurring supergene \textit{Segregation Distorter} (\textit{SD}), in order to investigate why \textit{SD} alleles remain relatively rare within populations despite being preferentially inherited. First, we show that the three \textit{SD} variants differ in the severity and dominance of the fitness costs they impose on individuals carrying them. Second, \textit{SD}-carrying parents produced less fit offspring in some crosses, independent of offspring genotype, indicating that \textit{SD} alleles can have non-genetic, transgenerational costs in addition to their direct costs. Third, we found that \textit{SD} carriers sometimes produce a biased offspring sex ratio, perhaps due to off-target effects of \textit{SD} on the sex chromosomes. Finally, we used a theoretical model to investigate how sex ratio and transgenerational effects alter the population genetics of distorter alleles; accounting for these additional costs helps to explain why real-world segregation distorter alleles are rarer than predicted by simpler models.}
---

<!-- author: -->
<!-- - Heidi W.S. Wong\textsuperscript{1} and Luke Holman*\textsuperscript{1} -->
<!-- - '*\textit{luke.holman@unimelb.edu.au} \vspace{5mm}' -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(xtable)
library(kableExtra)
library(stringr)
options(xtable.comment = FALSE)

table1 <- read.csv("../figures/Table1.csv")
table1$Difference <- gsub(" ", "", table1$Difference)
table1$Difference <- gsub("to", " to ", table1$Difference)
table1$Difference <- gsub("\\(", " \\(", table1$Difference)
table1$Comparison <- gsub("-", "→", table1$Comparison)
names(table1)[names(table1) == "Difference"] <- "Absolute diff."
names(table1)[names(table1) == "Relative.difference"] <- "Relative diff."
table1 <- table1[, !(names(table1) %in% "Error")]
table1$Implication <- NA
table1$Comparison <- gsub(" parents", "", table1$Comparison)
table1$Implication[table1$Comparison == "Both, 0 or 1 → Both, 2"] <- "Extra costs when homozygous"
table1$Implication[table1$Comparison == "Both, 1 → Both, 2"] <- "Extra costs when homozygous"
table1$Implication[table1$Comparison == "Neither, 0 → Mother, 0"] <- "Transgenerational effect"
table1$Implication[table1$Comparison == "Mother, 0 → Father, 0"] <- "Transgenerational effect"
table1$Implication[table1$Comparison == "Mother, 0 → Father, 0"] <- "Transgenerational effect"
table1$Implication[table1$Comparison == "Mother, 1 → Father, 1"] <- "Transgenerational effect"
table1$Implication[table1$Comparison == "Neither, 0 → Father, 0"] <- "Transgenerational effect"
table1$Implication[table1$Comparison == "Mother, 0 → Mother, 1"] <- "Costs of SD to heterozygotes"
table1$Implication[table1$Comparison == "Father, 0 → Father, 1"] <- "Costs of SD to heterozygotes"
table1$Implication[table1$Comparison == "Father, 0 → Father, 1" & table1$SD == "SD-Mad"] <- "Benefits of SD to heterozygotes"

# Negative number means the first group is lower than the second
table2 <- read.csv("../figures/Table2.csv")
table2$Difference <- gsub(" ", "", table2$Difference)
table2$Difference <- gsub("to", " to ", table2$Difference)
table2$Difference <- gsub("\\(", " \\(", table2$Difference)
table2$Comparison <- gsub("-", "→", table2$Comparison)
names(table2)[names(table2) == "Difference"] <- "Absolute diff."
names(table2)[names(table2) == "Relative_difference"] <- "Relative diff."
table2 <- table2[, !(names(table2) %in% "Error")]
table2$Implication <- NA
table2$Implication[table2$Comparison == "Daughters, CyO, mother → Daughters, SD, mother"] <- "SD lowers survival"
table2$Implication[table2$Comparison == "Sons, CyO, mother → Daughters, CyO, mother"] <- "Sons have lower survival"
table2$Implication[table2$Comparison == "Sons, CyO, mother → Sons, CyO, father"] <- "Transgenerational effect"
table2$Implication[table2$Comparison == "Sons, SD, mother → Daughters, SD, mother"] <- "Sons have lower survival"
table2$Implication[table2$Comparison == "Sons, SD, mother → Sons, SD, father"] <- "Transgenerational effect"
table2$Implication[table2$Comparison == "Sons, CyO, father → Daughters, CyO, father"] <- "Sons have lower survival"

table1 <- xtable(table1, label="tb:expt1", 
                 size = "footnotesize", digits = c(0,0,0,0,0,0,3,0),
                 caption = "\\footnotesize{List of all the notable differences between groups in Experiment 1 (posterior probability, $p$, $<$0.05; see Tables S6-S9 for results that did not meet this arbitrary cutoff). For each contast, we list the parent(s) that carried \\textit{SD} (neither, mother, father, or both) and the number of \\textit{SD} alleles carried by the offspring. The absolute difference in means is expressed in the original units (i.e. \\% larvae surviving, \\% sons, per-female progeny production, or \\% offspring sired), and the parentheses give its 95\\% credible intervals. The difference is positive when the second-listed mean is higher than the first one, and negative otherwise (e.g. the first row indicates that the fitness of females carrying 0 copies of \\textit{SD-5} is higher when the mother rather than the father carries \\textit{SD-5}). The Relative difference column shows the second-listed mean divided by the first-listed one, and thus gives the difference as a ratio. The posterior probability $p$ is the chance that this difference is actually zero or has the opposite sign, given the priors, model, and data. The final column gives a biological interpretation for each difference; note that our experimental design cannot distinguish genomic imprinting from parental effects that differentially affect \\textit{SD} and non-\\textit{SD} offspring.}") 

                 
table2 <- xtable(table2, label="tb:expt2", 
                 size = "footnotesize", digits = c(0,0,0,0,0,3,0),
                 caption = "\\footnotesize{List of the all the notable differences between groups in Experiment 2 (posterior probability, $p$, $<$0.05; see Table S10 for results that did not meet this arbitrary cutoff). For each group, we list the sex of the focal larvae, their genotype (\\textit{SD} or \\textit{CyO}), and the parent that carried \\textit{SD} (mother or father). The difference in means is expressed in \\% larvae surviving; other details are as in Table 1.}") 

add_significance_stars <- function(tabl){
  if("Posterior probability" %in% names(tabl)){
    tabl$sig <- " "
    tabl$sig[tabl$`Posterior probability` < 0.05] <- "*"
    names(tabl)[names(tabl) == "sig"] <- " "
  }
  tabl
}

make_table <- function(file, digits = 3){
  
  tabl <- add_significance_stars(readRDS(file))
  if("Comparison" %in% names(tabl)) tabl$Comparison <- gsub("-", "→", tabl$Comparison)

  kable(tabl, "latex", booktabs = TRUE, digits = digits) %>%
    kable_styling(latex_options = c("repeat_header", "scale_down"), font_size = 8)
}
```

<!-- \maketitle\unmarkedfntext{\textsuperscript{1}School of BioSciences, The University of Melbourne, Victoria, Australia.} -->

\textbf{Keywords: gene drive, meiotic drive, population genetic model, selfish genes, \textit{t} paradox}.


\newpage
# Introduction
Segregation distorters are genetic elements that manipulate meiosis or gametogenesis such that they are present in more than the usual 50% of the gametes [@burt2006; @lindholm2016]. Because of this bias in transmission, segregation distorters are predicted to spread rapidly to fixation assuming that individuals carrying the distorter are equally fit as non-carriers [@bruck1957]. Even if a distorter allele reduces the fitness of individuals that carry it, it can still be favoured by selection provided that its individual-level fitness costs are outweighed by the within-individual advantage conferred by segregation distortion [@lindholm2016]. For this reason, there is currently great interest in using natural or artificially-created segregation distorters to spread human-beneficial alleles through wild populations, for example to introduce malaria resistance alleles into wild mosquito populations [@gantz2015]. In addition to their promise for applied science, the study of segregation distorters has led to multiple advances in our understanding of evolution, genetics, and speciation [@rice2013; @lindholm2016; @manser2017; @lin2018; @verspoor2018]. The best-studied naturally-occurring distorters are the _t_ allele in mice [@carroll2007], and the _Segregation Distorter_ (_SD_) allele of _Drosophila melanogaster_ [@larracuente2012], both of which caused biased tranmission in heterozygous males by preventing the development of sperm that did not inherit the distorter. 

The '_t_ paradox' [@carroll2007] is a long-standing evolutionary puzzle. Though the paradox is named after the mouse _t_ allele, it describes a puzzle that applies to many other segregation distorters that have similar properties [reviewed in @lindholm2016]. The paradox is that many distorter alleles are quite rare within populations despite their strong transmission advantage. For example, the _t_ allele occurs at frequencies of 5-14% depending on the population [@ardlie1998], and _SD_ occurs at frequencies of 0-8% [@brand2015], both of which are substantially lower than predicted by simple population genetic models [@bruck1957; @lewontin1968; @charlesworth1978; @taylor2002; @holman2015]. Taking the _t_ allele as an example, we know that _t_ is transmitted to a fraction $k$ of the offspring of heterozygous males where $k \approx 0.95$, and that individuals homozygous for _t_ generally have close to zero fitness [@ardlie1998]. Assuming no other effects on fitness or inheritance, a distorter like $t$ is predicted to reach an equilibrium allele frequency of $\frac{1}{2} - \sqrt{(k(1-k))} / 2k$ [@bruck1957], which is 38.5% for $k$ = 0.95. The discrepancy between this prediction and real-world allele frequencies indicates that something is missing from the model, and so several subsequent models sought to resolve the puzzle by incorporating additional biological details. For example, Lewontin [-@lewontin1962] argued that population structure can reduce the equilibrium frequency of a distorter allele [see also @bull2019], and Lewontin [-@lewontin1968] showed that strong, partially recessive fitness costs can reduce the equilibrium frequency of the distorter allele. Additionally, males carrying segregation distorters often perform worse in sperm competition due to the loss of half their sperm, which can affect evolution of the distorter allele under certain conditions [@taylor2002; @holman2015; @lindholm2016]. 

Here, we attempt to explain the puzzling rarity of _Segregation Distorter_ (_SD_) in _D. melanogaster_. Similar to _t_ in mice, _SD_ is a gene complex or 'supergene' [@thompson2014] composed of several linked loci on an autosome (chromosome 2). _SD_ causes strong segregation distortion in heterozygous males by disrupting the development of non-_SD_-carrying spermatids [reviewed in @larracuente2012]. The _SD_ supergene contains an 'insensitive' allele at the _Responder_ locus (_Rsp_), while most chromosomes that lack _SD_ carry a 'sensitive' _Rsp_ allele that makes them susceptible to distortion. Chromosomal inversions in the _SD_ region help to keep the component loci in linkage by suppressing recombination, which prevents the creation of recombinant 'suicide chromosomes' in which the insensitive _Rsp_ allele linked to _SD_ is replaced by a sensitive _Rsp_ allele. The threat of suicide chromosomes appears to have selected for reduced recombination, and the small number of loci that cause segregation distortion are usually embedded in a large non-recombining region comprising \textit{c}. 10% of the entire genome [@presgraves2009], which has accumulated deleterious mutations that have hitchhiked along with the distorter [@temin1984; @larracuente2012; @brand2015]. All _SD_ alleles are thought to descend from a single common ancestor that appeared around 38,000 years ago [@brand2015], though _SD_ has since diversified into multiple variants that differ in their inversions and in their load of deleterious mutations [@larracuente2012; @presgraves2009; @brand2015]. In some populations, _SD_ chromosomes are present at low, stable frequencies that suggest balancing selection [e.g. 0-8% in 14 populations; @brand2015], although high and unstable allele frequencies have also been reported: one _SD_ variant increased in frequency from 17% to 98% over 23 years in Wisconsin [@temin1984].

The evolutionary dynamics of distorters such as _SD_ depend strongly on the fitness of drive-carrying individuals [e.g. @lewontin1968]. Negative frequency-dependent selection is of particular interest, because it can maintain a balanced polmorphism of distorting and non-distorting alleles. If selection on the distorter is not negatively frequency-dependent, the distorter will eventually fix or go extinct [@holman2015]. Recessive fitness costs are one likely source of negative frequency-dependent selection, because recessive costs are expressed more often when the distorter allele (and thus distorter homozygotes) is common. However, some distorter alleles have no obvious fitness cost [@temin1984; @price2012], meaning that recessive costs probably cannot provide a complete answer to the _t_ paradox. Additionally, models [e.g. @bruck1957; @lewontin1968] demonstrate that homozygote lethality alone is insufficient to explain the low allele frequencies of strong distorters like _SD_ or _t_. For these two reasons, we also tested whether _SD_ has fitness costs besides being harmful when homozygous.

Here, we focus on the three best-studied variants of _SD_, which are named _SD-5_, _SD-72_, and _SD-Mad_ [all originally collected in Wisconsin; @sandler1959]. _SD-5_ carries a different set of inversions than the other two, and is thought to be homozygous lethal [@larracuente2012], while some _SD-72_- and _SD-Mad_-type alleles are reportedly fit as homozygotes [@temin1984]. Indeed, the _SD-Mad_ allele studied here was previously reported to be fully viable and fertile in both sexes when homozygous [@brittnacher1983co], making it especially puzzling that this _SD_ variant is not more common. To our knowledge, the relative fitness of _SD_ heterozygotes has never been measured, and homozygotes have only been scored as viable or non-viable; we thus sought to measure the three genotypes' relative fitnesses, which are crucial to the evolutionary dynamics of _SD_ [@lewontin1968]. We measured the fitness of each _SD_ genotype in the juvenile stage, as well as in male and female adults. We also investigated older reports [@hiraizumi1967; @denell1969] that the offspring sex ratio of males carrying _SD_ deviates from the usual 50:50. If autosomal distorter alleles like _SD_ alter the sex ratio in addition to their other effects, there would be presumably be evolutionary consequences [since there is strong, 'Fisherian' selection on the sex ratio; @fisher1930]. We therefore wrote a model to predict how sex ratio bias would affect allele frequencies of _SD_. Lastly, we tested whether _SD_ has non-genetic, transgenerational fitness effects, e.g. mediated by parental effects or genomic imprinting, and used a model to investigate how _SD_ evolves in the presence of such transgenerational effects. Our empirical and theoretical findings have implications for the evoluton of _SD_ and other natural and human-engineered distorter alleles, and help to resolve the _t_ paradox.


# Methods

## Fly stocks
All flies were reared at 25^o^C under natural light (_c._ 14h day length) in 25mm plastic vials containing food medium (yeast-soy-cornmeal-agar-corn syrup). All stocks were obtained from the Bloomington _Drosophila_ Stock Centre unless otherwise stated (_SD_ stock numbers: 64322, 64324, and 64323). 

In order to generate a non-_SD_ reference allele which also allowed us to visually distinguish flies carrying 0, 1 or 2 copies of _SD_, we created a stock carrying an isogenic copy of chromosome 2 that carried one recessive and one dominant 'marker' mutation. The recessive marker was a mutant allele of _bw_ encoding brown eye colour (obtained from a teaching laboratory in Melbourne; unknown origin), while the dominant marker was the transgene _Ubi-GFP_ (stock 5826), which expresses green fluorescent protein (GFP) throughout the body. To recombine these markers, we crossed F1 _bw/Ubi-GFP_ females to _bw_ males and collected male progeny expressing brown eyes and GFP. From these recombinants, we selected a single male and crossed it to a female carrying wild-type X chromosomes (one from the _bw_ stock and one from the _SD-72_ stock) as well as the balancer chromosome _SM5_, collected +/+; _bw-GFP/SM5_ progeny, and crossed them to create what we hereafter call the _bw-GFP_ stock.

In the adult fitness assays, we used opposite-sex _bw_ individuals as mates, and _Gla/CyO_ individuals (stock 44227) as same-sex competitors. The offspring of _Gla/CyO_ flies express a dominant mutant phenotype, distinguishing them from the offspring of the focal flies.

Lastly, the three SD-bearing Bloomington stocks had different balancer chromosomes (_SD-5_ used _CyO_, _SD-72_ used _SM5_, and _SD-Mad_ was not balanced), so we first re-balanced the three SD stocks to use the _CyO_ balancer (from the _Gla/CyO_ stock) to remove this potential confounding effect. We then crossed _SD/CyO_ progeny to the _bw-GFP_ stock to create _SD/bw-GFP_ individuals. 

## Pilot study: Confirming that _SD_ shows biased inheritance

We first ran a pilot to confirm that _SD_ is inherited by >50% of the adult progeny of heterozygote males. We mated 45 pairs, each consisting of a _bw_/_bw_ female and _SD_/_bw_ male, and recorded the sex and eye colour of each of the 4,016 resulting progeny (n = 16 crosses involved _SD-5_, 14 _SD-72_, and 15 _SD-Mad_). We then fit a binomial GLMM (with family as a random effect) to estimate the average % _SD_ progeny carrying _SD_ among the F1 sons and daughters reaching adulthood for each of the three _SD_ variants. Note that this method will underestimate the strength of segregation distortion if _SD_ progeny are more likely to die before reaching adulthood: it thus provides a lower bound on the proportion of _SD_-bearing sperm produced by heterozygote males.


## Experiment 1

### Experimental crosses
We performed four types of experimental crosses for each of the three _SD_ alleles (Figure 1). In Cross 1, we mated two _SD/bw-GFP_ flies, yielding offspring carrying 0, 1 or 2 _SD_ alleles. In Cross 2, we mated _SD/bw-GFP_ females to _bw_ males, yielding offspring carrying 0 or 1 _SD_ alleles. Cross 3 was the reciprocal of Cross 2: a _bw_ mother and _SD/bw-GFP_ father. Lastly, to measure the baseline fitness of the _bw-GFP_ allele in the same experimental conditions, we mated two _bw-GFP_ flies (Cross 4). 

All of these crosses were performed in parallel on a common cohort of flies under identical conditions in a randomised order, minimising confounding effects. We ran all four crosses (and their associated fitness assays; see below) in each of three experimental blocks, with equal representation of crosses within blocks. We measured three components of fitness: survival rate from first-instar larva (hereafter 'L1 larvae') to adult, adult male competitive fertilisation success, and adult female fecundity following social interaction. For brevity, we term these juvenile, male, and female fitness. We also recorded the adult sex ratio produced by each cross.

### Juvenile fitness and sex ratio assays
Mated females from the four experimental crosses were placed separately onto egg collection plates (grape-agar medium with live yeast) for 24h, then removed. We waited 24h, then collected L1 larvae and sorted them by GFP phenotype. The reason for beginning the assay with L1 larvae, not eggs, was that we could correctly classify the GFP phenotype of L1 larvae (100/100 successes in a pilot) but not eggs, and because it is difficult to distinguish unfertilised eggs from fertilised eggs in which the embryo died before hatching. We placed the sorted larvae in fresh vials in groups of up to 100. It was difficult to obtain 100 larvae for every class of progeny because some progeny classes are rare due to segregation distortion and/or pre-hatching mortality. We subsequently quantified juvenile fitness and the sex ratio by counting, sexing, and phenotyping the adults that eclosed from these vials. 

### Adult female and male fitness assays
Flies that survived to adulthood in the juvenile fitness assay were sorted by phenotype/genotype into single-sex vials, left to mature for 48-72h, and then used in adult fitness assays. 

To measure female fitness, we placed 5 same-genotype females in an 'interaction vial' with 15 _bw_ males and 10 _Gla/CyO_ females (all flies were 48- to 72-hour-old virgins), and allowed them to interact for 48h to facilitate mating, courtship, behavioural interactions, and competition for food. We then recorded the number of surviving focal females, and moved them as a group to a new yeasted food vial (without the non-focal flies), where they oviposited for 24h. We then removed the females and counted the number of larvae eclosing from their eggs, and used this as our measure of female fitness. Thus, our measure of female fitness measure is the product of female fecundity, the proportion of eggs that are fertilised, and offspring survival in the zygote-to-L1 stage. 

To measure male fitness, we placed 5 same-genotype males in an interaction vial with 15 _bw_ females and 10 _Gla/CyO_ males (again, all flies were 48- to 72-hour-old virgins), where they interacted and mated for 48h. We then moved all surviving individuals (focal and non-focal) to a new food vial where they continued to interact and oviposit for 24h. We then removed all adults and allowed their offspring to develop to adulthood, then counted the number of progeny sired by the focal males and the competitor _Gla/CyO_ males. We used the proportion of progeny sired by the focal males as a measure of adult male fitness. This fitness measure encompasses pre- and post-copulatory sexual selection, as well as the survival rate of focal males’ offspring relative to those of _Gla/CyO_ males. 

### Limitations of Experiment 1’s juvenile fitness assay
Upon phenotyping adult flies emerging from Crosses 1-4, we observed unexpected recombination between the _bw_ and _Ubi-GFP_ loci for the _SD-72_ and _SD-Mad_ (but not _SD-5_) chromosomes [we had assumed that _SD_ chromosomes would be largely non-recombining in light of previous data; e.g. @presgraves2009]. Specifically, in Cross 2, some GFP-negative larvae developed brown eyes, and some GFP-positive ones developed red eyes, indicating recombination in the _SD/bw-GFP_ mother (recombinants were never seen in Cross 3, because there is no recombination in male _Drosophila_; this shows that recombination rather than phenotyping errors explains the results). The proportion of recombinant adults in Cross 2 was 3.6% (95% CIs: 2.4-4.9%) for _SD-5_, 36.1% (33-39%) for _SD-72_, and 32.8% (30-36%) for _SD-Mad_. The _bw_ locus is at the terminal end of the right arm of chromosome 2 (2R), and _SD-5_ is distinguished from the other two variants by an additional inversion on 2R; we therefore hypothesise that the _Ubi-GFP_ transgenic insertion lies somewhere on 2R between the _SD_ complex and _bw_, probably close to the _SD-5_-specific inversion [Figure 1 in @larracuente2012]. As a consequence of this unexpected recombination, we cannot be certain how many larvae of each genotype were present at the start of the juvenile fitness assay for Cross 2, at least for _SD-72_ and _SD-Mad_ -- we simply removed the recombinant individuals from the dataset, and made the simplistic assumption that all of the larvae that did not reach adulthood were non-recombinants. We interpret the relevant part of the Results in light of the resulting bias. This limitation is offset by data from Experiment 2 (which does not rely on these markers, and uses a balancer chromosome to suppress recombination), as well as data from Cross 3 (since there is no recombination in male _Drosophila_).

Additionally, for Cross 1, individuals carrying 0 or 1 _SD_ chromosomes were phenotypically indistinguishable until they reached adulthood and developed eyes, and so we simply measured the survival rate of a mixed pool of larvae carrying either 0 or 1 _SD_ alleles. The great majority of larvae in this pool will carry 1 _SD_ allele, rather than 0, because of segregation distortion. Specifically, the proportion of _SD_ progeny in this pool will be $1 / (k + 2(1-k))$, or 95.2% for $k$ = 0.95. This limitation is offset by data from Crosses 2 and 3 and Experiment 2.

## Experiment 2
Experiment 2 was designed to measure the direct and transgenerational effects of _SD_ on sex-specific larval survival, and to address the limitations of Experiment 1. Experiment 2 used the transgenic construct _P{Sxl-Pe-EGFP.G}G78b_ (extracted from stock 24105, backcrossed into the $w^{1118}$ genotype for 5 generations, and made homozygous), which allows discrimination of males and females at the egg stage [female-destined embryos express GFP while males do not; @thompson2004]. We conducted six types of crosses using parents bred at standardised density: in each cross, one parent was _SD/CyO_ and the other was homozygous for _P{Sxl-Pe-EGFP.G}G78b_; we performed this cross with the three _SD_ variants, with either the mother or the father providing _SD_ (10-24 replicates per cross). We then collected embryos of both sexes (mean: 48 embryos per sex per cross), placed them in single-sex vials to develop, and then counted and phenotyped the eclosing adults to infer the survival rates of different progeny classes. 

## Statistical analysis
We analysed Experiment 1 using Bayesian hierarchical models implemented in the R package `brms` [@burkner2017]. The data on juvenile fitness, male fitness, and adult sex ratio were treated as binomially distributed, and we fit 'vial' as a random effect to account for nonindependence of measurements from the same vial (this random effect was unnecessary for the female fitness data, which had one observation per vial). Female fitness was modelled using the negative binomial distribution, since the data were overdispersed counts. For all fixed effects, we used a moderately informative prior (a normal distribution centered on zero with SD = 5), in order to regularise the parameter estimates and reduce overfitting [@mcelreath2018st]. We verified model fit using posterior predictive checks [@gelman2006]. 

For hypothesis testing, we calculated the posterior differences between pairs of means for contrasts that we deemed informative for this study. For example, we calculated the posterior difference between the mean fitnesses of individuals with 0 or 1 _SD_ allele, or individuals that received _SD_ from their father versus their mother, and thereby tested for genetic and parental effects respectively. We also calculated the posterior probability that the group with the larger posterior mean actually has a smaller mean than the other group; this provides a metric with a similar interpretation to the $p$-value (contrasts for which >95% of the posterior lies on one side of zero were considered notable). It is not necessary to correct for multiple testing when calculating these pairwise differences, since the contrasts are all calculated using the posterior from the same model and thus are not independent tests.

The aim of Experiment 2 is to estimate the proportion of _SD_ and non-_SD_ male and female larvae that survive to adulthood. However, because the genotype of larvae could not be visually determined at the start of Experiment 2, we had to estimate the initial numbers of larvae belonging to each genotype in order to calculate the survival rates of each genotype. For example, if we placed 50 larvae in a vial and 20 non-SD and 20 _SD_ individuals reached adulthood, we inferred the genotypes of the 10 dead ones. This unmeasured variable depends on the gametes produced by the _SD/CyO_ parent. Because _SD_ only causes distortion in males [@larracuente2012], we assumed that the _SD/CyO_ mothers transmitted _SD_ to 50% of their progeny. We also assumed 50% transmission in _SD/CyO_ fathers (i.e. $k$ = 0.5), in light of evidence that _CyO_ carries an insensitive allele of _Rsp_ that makes it immune to segregation distortion [@ganetzky1977]. We then used a binomial random number generator with $p$ = 0.5 to stochastically 'fill in' the genotypes of the dead larvae. Our sample size was sufficiently large that generating a new set of random numbers and re-running the model gave near-identical parameter estimates and identical qualitative conclusions, thanks to the law of large numbers. We also re-ran the model under the assumption that there is some segregation distortion in _SD/CyO_ fathers [i.e. \textit{k} > 0.5, contradicting the evidence in @ganetzky1977], and found that all the key results did not change (Figure S1).

## Population genetic model
Our experiments suggested that some _SD_ variants have parent-of-origin-specific effects on fitness, and that some _SD_ variantsc cause males to produce a biased offspring sex ratio. We therefore constructed a simple one-locus, two-allele population genetic model to examine the effect of these two factors on the evolution of _SD_. 

The model considers the spread of an autosomal segregation distorter in an infinitely large, panmictic population with discrete generations. We assume that individuals carrying two wild type alleles have a relative fitness of 1, while genotypes carrying _SD_ potentially have relative fitness betwen 0 and 1. We tracked the parental origin of the _SD_ allele in heterozygotes, to allow heterozygotes with a maternally-inherited _SD_ to have a different fitness than heterozygotes with a paternally-inherited _SD_, and thereby allow for parent-of-origin-specific effects on fitness. We assumed that male heterozygotes transmit _SD_ to a fraction $k$ offspring (where 0.5 < $k$ < 1), and produce a fraction $(1 + s)/2$ female offspring (-1 < $s$ < 1), while all other genotypes were assumed to show normal Mendelian inheritance and a 50:50 offspring sex ratio. 

For each parameter space, we determined the evolutionary fate of an _SD_ allele in a starting population with 1% _SD_ alleles at Hardy-Weinberg genotype frequencies. We calculated the equilibrium allele frequencies numerically, since the analytical solution would be unwieldy. In each generation, we first multiplied the frequency of each genotype by its relative fitness (representing the combined action of natural and sexual selection across all life stages) and then renormalised the genotype frequencies to sum to one. We then determined the frequency of each of the possible mating types as the product of each possible pair of maternal and paternal genotype frequencies. From these, we determined the offspring genotype frequencies, and replaced the parental generation with the offspring. The simulation ran for 10,000 generations to ensure that _SD_ had reached equilibrium, though it was terminated early if _SD_ went extinct (defined as reaching 0.001% frequency) or fixed (>99%).

# Results

## Biased inheritance of _SD_
We found that the percentage of adult progeny carrying _SD_ in crosses where the father was an _SD_ heterozygote was 88% (95% CIs: 85-90%) for _SD-5_, 89% (86-91%) for _SD-72_, and 82% (79-86%) for _SD-Mad_ (Figure S2). The percentage was significantly lower for _SD-Mad_ (p = 0.009), implying that this variant has weaker segregation distortion and/or better egg-to-adult survival. The proportion of _SD_ progeny was similar among the sons and daughters of _SD_ males, and there was no interation between _SD_ variant and offspring sex (all p > 0.084). 

## Experiment 1
Posterior estimates of mean fitness for each group are plotted in Figure 1. Tables S1-S4 give sample sizes and summary statistics, and Tables S5-S8 present estimated differences between means. Table 1 summarises Tables S5-S8 by listing only the differences for which >95% of the posterior lies on one side of zero.

\begin{figure}
\centering
\includegraphics{../figures/fig1_inkscape.pdf}
\caption{\footnotesize{Posterior estimates of the group means for the four different response variables in Experiment 1, for each type of cross ($x$-axis), \textit{SD} variant (panels), and offspring genotype (colours). Juvenile fitness was measured as \% L1 larva-to-adult survival, adult sex ratio refers to the number of males and females among the individuals that reached adulthood, female fitness is the estimated number of progeny produced per female, and male fitness is the siring success relative to competitior males. The thicker inner bar shows the region containing 50\% of the posterior, the outer bar covers 95\% of the posterior, and the circle marks the median. Tables S6-S9 give the accompanying statistical results. Points labelled as carrying ``0 or 1'' \textit{SD} allele refer to cases where the genotype of the offspring could not be ascertained; most of these individuals ($>90\%$) probably carried 1 \textit{SD} allele because of segregation distortion.}}
\end{figure}

```{r results='asis'}
print(table1, booktabs=TRUE, size='scriptsize', include.rownames = FALSE, caption.placement = "top")
```

### Juvenile fitness
When collecting larvae we observed 40 L1 larvae homozygous for _SD-5_, and over 600 carrying two copies of _SD-72_, but not one of these larvae survived to adulthood. Since we collected roughly the same number of eggs for each _SD_ variant (i.e. well over 600 eggs), the smaller number of _SD-5_ larvae indicates that most _SD-5_ homozygotes died before hatching, while _SD-72_ homozygotes primarily died between hatching and adulthood. By contrast, many larvae homozygous for _SD-Mad_ reached adulthood, and there was no statistically significant effect of _SD-Mad_ on larval survival, even in homozygotes (the % larvae surviving was lower by 10% among individuals carrying two copies of _SD-Mad_ rather than one, with 95% CIs of -6.7 to 27.7; Table S6).

The limitations of this assay (see Methods) mean that Figure 1 might underestimate the survival rate of individuals carrying a maternally-inherited _SD_ allele, for _SD-72_ and _SD-Mad_. Therefore, we cannot be certain that there is really no difference in juvenile fitness between individuals with an _SD_ mother versus an _SD_ father for _SD-72_ and _SD-Mad_.


### Sex ratio among individuals reaching adulthood
For crosses in which the father carried _SD-Mad_, the sex ratio of the emerging adults was significantly more female-biased than for crosses in which the mother carried _SD-Mad_, irrespective of offspring genotype. The results did not replicate earlier findings that the non-_SD_ offspring of _SD_ heterozygote fathers show a female-biased sex ratio [@hiraizumi1967; @denell1969]; indeed, there was a nonsignificant trend in the opposite direction for _SD-5_ (the posterior median was 54% sons among the non-_SD_ offspring, and 48% sons among the _SD_ offspring; Figure 1).

### Adult female fitness
Although _SD-Mad_ homozygotes were viable, female homozygotes produced far fewer progeny than female heterozygotes from the same cross. There was evidence that _SD-Mad_ had non-genetic transgenerational effects on female fitness: the non-_SD_ daughters of _SD_ mothers were fitter than non-_SD_ daughters whose _father_ carried _SD_. Indeed, the non-_SD_ daughters of _SD_ mothers were actually fitter than daughters from the control cross in which neither parent carried _SD_. The same results were found for _SD-5_: the non-_SD_ daughters of _SD-5_ mothers were more fit than those of _SD-5_ fathers, or daughters from the control cross. _SD-5_ also had a direct genetic effect on female fitness: females carrying _SD-5_ had lower fitness than females from the same cross that did not inherit it, although this effect was only observed when _SD-5_ was maternally inherited. _SD-72_ had no detectable effects on female fitness, other than causing complete lethality of juveniles when homozygous.

### Adult male fitness
Males homozygous for _SD-Mad_ had low fitness. We again observed evidence for non-genetic transgenerational effects: for _SD-5_, males with a paternally-inherited _SD_ chromosome were substantially less fit than males with a maternally-inherited _SD_ chromosome. There was also a strong fitness cost of inheriting an _SD-5_ allele from either parent, suggesting that _SD-5_ has a harmful direct genetic effect on male fitness. Interestingly, the sons of _SD-Mad_ fathers were fitter if they inherited _SD-Mad_ rather than the non-_SD_ allele; a similar though non-sigificant result (posterior probability = 0.080; Table S8) was observed in the female fitness assay. 

## Experiment 2
Experiment 2 suggested that _SD_ chromosomes can have both direct and transgenerational effects on L1 larva-to-adult survival (Figure 2; Table 2; full results in Tables S9-10). Male larvae with an _SD-5/CyO_ mother were significantly less likely to survive than those with an _SD-5/CyO_ father, irrespective of whether the larva actually inherited _SD-5_. A similar result was observed for _SD-Mad_, though only among offspring that inherited _CyO_ rather than _SD_. Also, for crosses in which the mother carried either _SD-5_ or _SD-Mad_, survival was lower among daughters that inherited _SD_ rather than _CyO_ (and since _CyO_ itself carries deleterious mutations, this implies that _SD_ would also lower fitness relative to the wild type). The same effect was not observed for male larvae, or for crosses in which _SD_ was inherited from the father, possibly indicating that _SD_ alleles can have sex- or parent-of-origin-specific effects on larval survival. Lastly, we observed some significant sex differences in survival for all three _SD_ chromosomes, with female larvae surviving better than male larvae for six different combinations of offspring and parental genotypes. We did not find any evidence that the direct genetic effect of _SD_ on larval survival is sex-specific: the (small) differences in survival between _SD_ and _CyO_ progeny were similar in sons and daughters (Figure 2). 


\begin{figure}
\centering
\includegraphics{../figures/fig2.pdf}
\caption{\footnotesize{Posterior estimates of \% L1 larva-to-adult survival in Experiment 2 for each combination of offspring sex and genotype ($x$-axis), \textit{SD} variant (panels), and whether the mother or father had the genotype \textit{SD}/\textit{CyO} (`cross'; colours). The thicker inner bar shows the region containing 50\% of the posterior, the outer bar covers 95\% of the posterior, and the circle marks the median. See Tables 2 and S10 for associated hypothesis tests. The model underlying this plot assumed fair meiosis ($k$ = 0.5) in \textit{SD/CyO} males; see Figure S1 for equivalent plots made using different assumed values of $k$.}}
\end{figure}

```{r results='asis'}
print(table2, booktabs=TRUE, size='scriptsize', include.rownames = FALSE, caption.placement = "top")
```


## Population genetic model
We first assumed that the _SD_ allele had no direct or transgenerational fitness costs (top left, Figure 3), which allowed _SD_ to invade even if segregation distortion ($K$) was very weak. However, if the _SD_ allele caused males carrying it to produce a highly biased sex ratio (unrealistically high, based on our data), _SD_ required a higher $K$ to invade. The reason that this sex ratio bias hinders the spread of _SD_ is that autosomal loci usually maximise their fitness by producing a 50:50 sex ratio, due to 'Fisherian' selection on the sex ratio, which disfavours alleles causing unequal production of sons and daughters [@fisher1930]. In cases where the _SD_ allele was able to invade, it generally went to fixation: a balanced polymorphism of _SD_- and non-_SD_ alleles was seldom observed. There was a small zone of polymorphism when drive was very weak and sex ratio bias was very strong (both of which are unrealistic for any known distorter alleles). This polymorphism results from the frequency-dependent selection on alleles that affect the sex ratio: over-producing one sex is especially costly if that sex is over-represented in the population.

Secondly, when we assumed that all individuals with at least one _SD_ allele had a relative fitness of 0.8 (dominant costs, top second left of Figure 3), the _SD_ allele could still invade, though it needed a substantially higher transmission bias $K$ to do so. When _SD_ could invade, it again proceeded to fixation, except under unrealistically weak drive and extreme sex ratio bias. Notably, invasion was more difficult (i.e. a higher $K$ was required) when we assumed that _SD_ heterozygote males produce a female-biased rather than male-biased sex ratio; this is because _SD_ can only bias segregation in males. _SD_ invaded slightly more easily when _SD_ heterozygote males produced >50% sons, but invasion was still harder than when _SD_ did not bias the sex ratio (due to Fisherian sex ratio selection against _SD_).

Thirdly, when we assumed that _SD_ is recessive-lethal but cost-free in heterozygotes (top second right, Figure 3), the _SD_ allele stabilised at high, intermediate frequencies for realistic (i.e. high) values of $K$ [as expected; @bruck1957]. This is because recessive fitness costs create negative frequency-dependent selection on _SD_, halting the spread of the _SD_ allele once homozygotes become common enough to cancel out the effect of segregation distortion [@holman2015]. A female-biased sex ratio reduced the equilibrium frequency of _SD_ while a male-biased sex ratio had little effect, due to the opposing effects of Fisherian selection and the benefits of producing more sons (i.e. the sex in which distortion occurs).

Fourthly, we modelled a recessive-lethal _SD_ that reduces the relative fitness of heterozygotes to 0.8 (top right, Figure 3 -- this assumption is probably the most realistic so far, based on our empirical findings). Here, the _SD_ allele only invaded when $K$ was high, and it stabilised at medium-high frequencies. Interestingly, _SD_ alleles that induced a male-biased sex ratio invaded for substantially lower $K$ and reached a higher equilibrium frequency for any given $K$ than those that did not affect the sex ratio. Presumably this occurred because when _SD_ is kept rare by its direct fitness costs, the population sex ratio stays close to 50:50, and so Fisherian sex ratio selection against _SD_ remains weak (while the benefits of extra transmission bias stay the same).

For all four of these scenarios, we produced similar graphs under the additional assumption that offspring suffer an extra cost when the _SD_ allele is inherited from a particular parent. In the middle row of Figure 3, genotypes carrying a paternally-inherited _SD_ allele have their fitness reduced by an additional 0.2, while in the bottom row, the same applies to genotypes with a maternally-inherited _SD_. Comparison of the three rows shows that these transgenerational costs further hamper the spread of _SD_, and that paternal costs are worse than maternal costs. The reason that paternal costs are worse is that they primarily afflict _SD_-carrying offspring (because of segregation distortion in males), while maternal costs harm a mixture of _SD_ and non-_SD_ offspring, reducing the impact of the transgenerational cost on the relative fitness of _SD_. By combining recessive lethality with some mixture of heterozygote fitness costs, sex ratio bias, or transgenerational costs, we could get _SD_ chromosomes to persist at low, stable frequencies as they often do in nature (e.g. the middle right panel of Figure 3 near $K$ = 0.95, which approximates the costs and $K$ value for _SD-5_). 


\begin{figure}[ht]
\centering
\includegraphics{../figures/fig3.pdf}
\caption{\footnotesize{The equilibrium frequency reached by the \textit{SD} allele depends on the strength of segregataion distortion ($K$ = 0 indicates fair meiosis, $K$ = 1 denotes complete distortion), as well as the direction and strength of sex ratio bias in the progeny of \textit{SD} heterozygote males. The four columns make different assumptions about the fitness costs to individuals carrying the \textit{SD} allele, while the three rows assume either that \textit{SD} has no parent-of-origin-specific effects on fitness (top row), or that \textit{SD} is especially costly when paternally inherited (middle row) or maternally inherited (bottom row). In the `Dominant costs' column and the fourth column, individuals with one copy of \textit{SD} had a relative fitness of 0.8, while `Recessive lethal' means that \textit{SD} homozygotes had zero fitness.}}
\end{figure}



# Discussion
Our results reaffirmed that _SD-5_ and _SD-72_ are homozygous lethal. Most _SD-5_ homozygotes died in the egg stage, while _SD-72_ homozygotes died after hatching but before adulthood. Although populations of _SD-Mad_ homozygotes can be cultured in the lab, and most of them survived until adulthood in Experiment 1, we found that adult _SD-Mad_ homozygotes had far lower fecundity and siring success than the comparison genotype (which was an inbred lab strain carrying two visible mutations). Thus, it seems plausible that _SD-Mad_ homozygotes might have close to zero fitness in the wild. The fitness costs to female and male adults were dominant for _SD-5_ but recessive for _SD-72_ and _SD-Mad_, suggesting that _SD-5_ carries additional dominant mutations that the others lack. Although we did not observe any _SD_ variants that had high fitness as homozygotes, it is possible that such variants exist; an _SD_ variant with inversions characteristic of _SD-72_ or _SD-Mad_ was reportedly present in 98% of individuals in a population in Wisconsin [@temin1984]. We also conducted a pilot study, and found that _SD_ is present in around 80-85% of the adult progeny of _SD_ heterozygote males; this is likely to be an underestimate of the strength of segregation distortion. 

Interestingly, we found some evidence for costly non-genetic transgenerational effects associated with _SD-5_ and _SD-Mad_. These transgenerational effects might represent parental effects [i.e. non-genetic effects of parental phenotype on offspring phenotype; @badyaev2009], genomic imprinting [i.e. when the effect of a genotype depends on the parental origins of the alleles; @holman2014], or a combination of both. Firstly, fitness was reduced among the non-_SD_ daughters of _SD-5_ or _SD-Mad_ heterozygote fathers, relative to heterozygote mothers. One possible mechanism is that non-_SD_-carrying chromosomes that escape segregation distortion are epigenetically modified in ways that affect adult fitness; this mechanism is plausible because _SD_ is thought to function by altering the chromatin of sensitive chromosomes [@larracuente2012]. Secondly, _SD-5_ was especially harmful to adult male fitness when paternally inherited, hinting at either genomic imprinting or a paternal effect of _SD-5_ that varies based on offspring genotype. Thirdly, in Experiment 2, we found that male larvae were less likely to reach adulthood when their mother carried _SD-5_ than when their father did, irrespective of whether the larva actually inherited _SD-5_. This result again suggests that _SD-5_ has a transgenerational effect on offspring fitness, though puzzlingly the harmful effect was associated with mothers rather than fathers this time (likely because Experiments 1 and 2 used a different non-_SD_ reference chromosome and genetic background). To our knowledge, all previous theoretical models of segregation distorters implicitly assume that transgenerational effects are absent. We therefore allowed _SD_ alleles to have parent-of-origin-specific effects on fitness in our model, revealing that such costs can reduce the invasion probability and equilibrium frequency of _SD_. Thus, if segregation distorters commonly have harmful transgenerational effects in addition to their direct cost to the individual carrying them, transgenerational costs may help to explain the puzzlingly low allele frequencies of _SD_ [@brand2015] and other autosomal distorters such as the _t_-haplotype [@carroll2007].

We also observed that fathers heterozygous for _SD-Mad_ produced an excess of daughters, while _SD-5_ and _SD-72_ parents produced a similar sex ratio to controls. Our results thus differ from earlier studies of _SD-5_ and _SD-72_, which found an excess of daughters but only among the non-_SD_ progeny [@hiraizumi1967; @denell1969]. In light of those earlier results, Larracuente and Presgraves [-@larracuente2012] proposed that Y-bearing spermatids might be eliminated in _SD_ males as a result of 'collateral damage' arising because of sequence homology between Y-linked loci and _Responder_, which could explain the observed shortage of sons in crosses where the father carries _SD_. As an alternative or complement to this hypothesis, we speculate that _SD_ might cause a parental effect that differentially affects the survival of sons and daughters, for example by inducing epigenetic modifications that are more harmful in males (this hypothesis was not supported by Experiment 2, but it was not definitively ruled out either). Our modelling results suggest that _SD_ alleles invade less easily, and reach a lower equilibrium frequency, when they cause male heterozygotes to produce a female-biased sex ratio. There are two reasons for this result: firstly, autosomal alleles that skew the sex ratio away from 50:50 are usually disfavoured by selection [@fisher1930], and secondly, _SD_ alleles can only distort segregation in sons. The model also showed that producing a male-biased sex ratio was disadvantageous for _SD_ alleles, except in populations where _SD_ was kept rare by its fitness costs. When _SD_ is rare, the population-wide sex ratio remains close to 50:50, reducing the Fisherian cost to _SD_ of producing extra sons. Assuming that other autosomal segregation distorters also cause imbalanced sex ratios, this finding may be relevant to resolving the _t_-paradox for other species' distorter alleles.

In a somewhat unexpected result, we found that the adult sons and daughters of _SD-Mad_-bearing fathers were fitter if they inherited _SD-Mad_, relative to those that did not inherit it. We also found that the larvae of _SD-Mad_-bearing fathers were more likely to survive until adulthood if they inherited _SD-Mad_ rather than the alternative _CyO_ chromosome in Experiment 2. Assuming these results are genuine and not statistical flukes, we can infer either that _SD-Mad_ heterozygotes were fitter than both _SD_-free test genotypes, or that _SD-Mad_ has transgenerational effects when transmitted by fathers. The _SD_ allele is thought to inactivate non-_SD_-bearing spermatids by altering their chromatin, and so it is possible that the few non-_SD_ gametes that do survive being inactivated carry epigenetic 'scars'. Assuming that sperm that escape segregation distortion produce lower-fitness progeny, we predict that _SD_ alleles will reach slightly higher equilibrium alleles frequencies than they otherwise would, since only non-_SD_ alleles would be harmed in this way.

Future studies could compete _SD_ alleles with differing costs, and differing cost dominance, in population cages. We predict that _SD_ alleles with dominant costs will either fail to spread (if the costs are sufficiently high relative to the strength of segregation distortion, $k$), or will sweep to fixation, while alleles with recessive costs will potentially reach an evolutionary equilibrium. Similarly, we predict that the stability and allele frequencies of _SD_ chromosomes in natural populations will correlate with their fitness costs in homozygotes and heterozygotes. In line with this prediction, _SD-5_ is more costly, has more dominant costs, and was rarer than other the other two variants in the original Wisonsin population [@temin1984], and it would be interesting to see if the frequencies of competing _SD_ variants can be similarly explained in other populations. Our results also have implications for the design of artificial gene drives, or attempts to use natural gene drives like $t$ to deliver human-beneficial 'payloads' [e.g. there are proposals to modify the _t_ allele to control invasive populations of mice; @backus2016ge]. We suggest considering the fitness of drive-carrying individuals' offspring (not just the fitness of the carriers themselves) when testing a newly-designed gene drive in the lab, since our model shows that transgenerational costs can strongly influence the invasion success of the gene drive. 


# Acknowledgements
We are grateful to Tom AR Price for helpful comments on the manuscript.

## Availability of data and code
All raw data and R code is available at [https://lukeholman.github.io/fitnessCostSD/](https://lukeholman.github.io/fitnessCostSD/).

# References
<div id="refs"></div>

\newpage
\pagenumbering{gobble}

\begin{center}
\huge{\textbf{Online Supplementary Material}}
\end{center}

\begin{center}
\textbf{Fitness consequences of the selfish supergene \textit{Segregation Distorter}}
\end{center}

\vspace{1.4cm}


The figures and tables in this document, along with the with the R code used to generate them, can also be viewed online: \newline \textbf{https://lukeholman.github.io/fitnessCostSD}


\newpage


# Supplementary figures

![](../output/fig_S1.pdf)

**Supplementary Figure 1**: Equivalent plots to Figure 2, under the assumption that meiosis is fair ($k$ = 0.5, top row, same as Figure 1), slightly biased ($k$ = 0.6, middle row), and more strongly biased ($k$ = 0.7, bottom row) in _SD_/_CyO_ males. Note that the significant results for Figure 2 mostly stay the same or increase in magnitude, suggesting that the results are not strongly influenced by our assumptions about the strength of segregation distortion in _SD_/_CyO_ males.
\newpage

\begin{figure}[ht]
\centering
\includegraphics{../output/fig_S2.pdf}
\end{figure}
**Supplementary Figure 2**: Posterior estimates of the percentage of adult progeny that carried _SD_ in the pilot experiment, which crossed _SD_/_bw_ males and _bw_/_bw_ females. All estimates lie well above the 50% expected under Mendelian inheritance, and the estimate for _SD-Mad_ is significantly lower than the other two. Note that the elevated percentage of _SD_ progeny is the net result of segregation distortion and pre-adult mortality, and so the strength of segregation distortion might be stronger than suggested by these estimates if _SD_ progeny are less likely to survive to adulthood.
\newpage


# Supplementary tables

**Supplementary Table 1**: Number and percentage of L1 larvae surviving to adulthood for each _SD_ genotype and cross type.
```{r, results='asis'}
make_table("../output/tab_S1.rds")
```
\newpage

**Supplementary Table 2**: Number and percentage of male and female adults emerging from the juvenile fitness assay vials.
```{r, results='asis'}
make_table("../output/tab_S2.rds")
```
\newpage

**Supplementary Table 3**: Average fecundity of adult females for each SD genotype and cross type. The last two columns give the sample size in terms of number of oviposition vials (each of which contained up to 5 focal females), and number of males.
```{r, results='asis'}
make_table("../output/tab_S3.rds")
```
\newpage

**Supplementary Table 4**: Average relative fitness of adult males for each SD genotype and cross type, expressed as the average proportion of offspring sired. The last two columns give the sample size in terms of number of vials (each of which contained 5 focal males), and number of males.
```{r, results='asis'}
make_table("../output/tab_S4.rds")
```
\newpage

**Supplementary Table 5**: The results of hypothesis tests computed using the model of larval survival in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if the first mean is larger than the second, and negative otherwise (expressed in % larval survival). The error column gives the average absolute residual, and other details are as in Tables 1 and 2.
```{r, results='asis'}
make_table("../output/tab_S5.rds")
```
\newpage

**Supplementary Table 6**: The results of hypothesis tests computed using the model of adult sex ratio in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if the first mean is larger than the second, and negative otherwise (expressed in % males). The error column gives the average absolute residual, and other details are as in Tables 1 and 2.
```{r, results='asis'}
make_table("../output/tab_S6.rds")
```
\newpage

**Supplementary Table 7**: The results of hypothesis tests computed using the model of female fitness in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if the first mean is larger than the second, and negative otherwise (expressed as the number of offspring produced). The error column gives the average absolute residual, and other details are as in Tables 1 and 2.
```{r, results='asis'}
make_table("../output/tab_S7.rds")
```
\newpage

**Supplementary Table 8**: The results of hypothesis tests computed using the model of male fitness in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if the first mean is larger than the second, and negative otherwise (expressed in % offspring sired). The error column gives the average absolute residual, and other details are as in Tables 1 and 2.
```{r, results='asis'}
make_table("../output/tab_S8.rds")
```
\newpage

**Supplementary Table 9**: Number and percentage of L1 larvae surviving to adulthood in Experiment 2, for each SD genotype, cross type, and offspring sex.
```{r, results='asis'}
make_table("../output/tab_S9.rds")
```
\newpage

**Supplementary Table 10**: Complete version of Table 2, showing all the contrasts that were tested in Experiment 2.
```{r, results='asis'}
make_table("../output/tab_S10.rds")
```

