---
title: "Supplementary Information"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{booktabs} \usepackage{microtype} \usepackage{longtable}
  \usepackage[margin=1in]{geometry}
---

## Fitness consequences of the selfish supergene _Segregation Distorter_

## Supplementary Figure and Supplementary Tables

All of the figures and tables in this document can also be viewed online at **https://lukeholman.github.io/fitnessCostSD/statistics.html**, along with the R code used to generate them. 

\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(kableExtra)
library(stringr)

make_table <- function(file, digits = 3){
  kable(readRDS(file), "latex", booktabs = TRUE, digits = digits) %>%
    kable_styling(latex_options = c("repeat_header", "scale_down"), font_size = 8)
}
```

# Supplementary figures


![](../output/fig_S1.pdf)

**Supplementary Figure 1**: Equivalent plots to Figure 2, under the assumption that meiosis is fair ($k$ = 0.5, top row, same as Figure 2), slightly biased ($k$ = 0.6, middle row), and more strongly biased ($k$ = 0.7, bottom row). Note that the significant results for Figure 2 mostly stay the same or increase in magnitude, suggesting that they are genuine and are not sensitive to our assumptions about the data.


\newpage
# Supplementary tables

**Supplementary Table 1**: Number and percentage of L1 larvae surviving to adulthood for each SD genotype and cross type.
```{r, results='asis'}
make_table("../output/tab_S1.rds")
```
\newpage

**Supplementary Table 2**: Number and percentage of male and female adults emerging from the juvenile fitness assay vials.
```{r, results='asis'}
make_table("../output/tab_S2.rds")
```
\newpage

**Supplementary Table 3**: Average relative fitness of adult males for each SD genotype and cross type, expressed as the average proportion of offspring sired. The last two columns give the sample size in terms of number of vials (each of which contained 5 focal males), and number of males.
```{r, results='asis'}
make_table("../output/tab_S3.rds")
```
\newpage

**Supplementary Table 4**: Average fecundity of adult females for each SD genotype and cross type. The last two columns give the sample size in terms of number of oviposition vials (each of which contained up to 5 focal females), and number of males.
```{r, results='asis'}
make_table("../output/tab_S4.rds")
```
\newpage

**Supplementary Table 5**: Number and percentage of L1 larvae surviving to adulthood in Experiment 2, for each SD genotype, cross type, and offspring sex.
```{r, results='asis'}
make_table("../output/tab_S5.rds")
```
\newpage

**Supplementary Table 6**: The results of hypothesis tests computed using the model of larval survival in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if mean 1 is larger than mean 2, and negative otherwise (expressed in % larval survival). The mean 1 and mean 2 columns list the parent which had SD (mother, father, or both), followed by the number of SD alleles present in the offspring (0, 1 or 2). The Posterior probability column gives the probability that the mean with the smaller point estimate is actually larger than the other mean, analagously to a one-tailed p-value. The Evidence ratio (ER) column gives the ratio of evidence, such that ER = 5 means that it is 5 times more likely that the mean with the smaller point estimate really is the smaller one. Asterisks highlight rows where the posterior probability is less than 0.05.
```{r, results='asis'}
make_table("../output/tab_S6.rds")
```
\newpage

**Supplementary Table 7**: The results of hypothesis tests computed using the model of adult sex ratio in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if mean 1 is larger than mean 2, and negative otherwise (expressed in % males). The mean 1 and mean 2 columns list the parent which had SD (mother, father, or both), followed by the number of SD alleles present in the offspring (0, 1 or 2). The Posterior probability column gives the probability that the mean with the smaller point estimate is actually larger than the other mean, analagously to a one-tailed p-value. The Evidence ratio (ER) column gives the ratio of evidence, such that ER = 5 means that it is 5 times more likely that the mean with the smaller point estimate really is the smaller one. Asterisks highlight rows where the posterior probability is less than 0.05.
```{r, results='asis'}
make_table("../output/tab_S7.rds")
```
\newpage

**Supplementary Table 8**: The results of hypothesis tests computed using the model of female fitness in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if mean 1 is larger than mean 2, and negative otherwise (expressed as the number of offspring produced). The mean 1 and mean 2 columns list the parent which had SD (mother, father, or both), followed by the number of SD alleles present in the offspring (0, 1 or 2). The Posterior probability column gives the probability that the mean with the smaller point estimate is actually larger than the other mean, analagously to a one-tailed p-value. The Evidence ratio (ER) column gives the ratio of evidence, such that ER = 5 means that it is 5 times more likely that the mean with the smaller point estimate really is the smaller one. Asterisks highlight rows where the posterior probability is less than 0.05.
```{r, results='asis'}
make_table("../output/tab_S8.rds")
```
\newpage

**Supplementary Table 9**: The results of hypothesis tests computed using the model of male fitness in Experiment 1. Each row gives the posterior estimate of a difference in means, such that the estimate is positive if mean 1 is larger than mean 2, and negative otherwise (expressed in % offspring sired). The mean 1 and mean 2 columns list the parent which had SD (mother, father, or both), followed by the number of SD alleles present in the offspring (0, 1 or 2). The Posterior probability column gives the probability that the mean with the smaller point estimate is actually larger than the other mean, analagously to a one-tailed p-value. The Evidence ratio (ER) column gives the ratio of evidence, such that ER = 5 means that it is 5 times more likely that the mean with the smaller point estimate really is the smaller one. Asterisks highlight rows where the posterior probability is less than 0.05.
```{r, results='asis'}
make_table("../output/tab_S9.rds")
```
\newpage

**Supplementary Table 10**: Complete version of Table 2, showing all the contrasts that were tested in Experiment 2.
```{r, results='asis'}
make_table("../output/tab_S10.rds")
```
