---
title: "Tables of the raw data"
date: "2019-03-12"
output: workflowr::wflow_html
---

```{r setup, cache=FALSE, include=FALSE}
library(knitr)
library(kableExtra)
library(dplyr)
opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```


```{r}
larvae  <- read.csv("data/clean_data/larval_fitness.csv") %>% 
  select(-SD, -MotherHasSD, -FatherHasSD, -SD_from_mum, -SD_from_dad, -parent_with_SD, -vial, -total, -larval_density) %>%
  select(mum_genotype, dad_genotype, copies_of_SD, block, number_larvae_in_vial, initial_number, everything()) %>%
  rename(initial_number_focal_larvae = initial_number)
males   <- read.csv("data/clean_data/male_fitness.csv") %>% 
  select(-SD, -MotherHasSD, -FatherHasSD, -SD_from_mum, -SD_from_dad, -parent_with_SD, -vial, -total, -focal.male.fitness, -rival.male.fitness) %>%
  select(mum_genotype, dad_genotype, copies_of_SD, block, everything())
females <- read.csv("data/clean_data/female_fitness.csv") %>% 
  select(-SD, -MotherHasSD, -FatherHasSD, -SD_from_mum, -SD_from_dad, -parent_with_SD, -vial) %>%
  select(mum_genotype, dad_genotype, copies_of_SD, block, everything())
expt2   <- read.csv("data/clean_data/experiment2.csv") %>%
  select(SD, SD_parent, sex_of_embryos, block, everything()) %>%
  select(-vial, -SD_embryos, -CyO_embryos) %>%
  rename(initial_number_larvae = starting_number_focal_sex) %>%
  rename(surviving_SD_offspring = SD_offspring) %>%
  rename(surviving_CyO_offspring = CyO_offspring) %>%
  arrange(SD, SD_parent, sex_of_embryos, block)
```

# Raw data from Experiment 1: larval survival and sex ratio

 - `mum_genotype`: The genotype of the mother (note that 'wild type' actually just means not SD; it was actually the _bw-GFP_ tester genotype, not true wild type).
 - `dad_genotype`: The genotype of the father.
 - `copies_of_SD`: The number of copies of SD carried by the focal individuals.
 - `block`: The experimental block. Flies from the same block were bred and measured concurrently.
 - `number_larvae_in_vial`: The total number of larvae in the vial (i.e. focal individuals, plus occasionally some _Gla/CyO_ larvae added to make up the numbers for very rare progeny classes)
 - `initial_number_focal_larvae`:The total number of larvae in the vial, excluding any _Gla/CyO_ larvae.
 - `surviving_females`: The number of females that eclosed as adults.
 - `surviving_males`: The number of males that eclosed as adults.
 - `survivors`: The number of males plus females that eclosed as adults.
 - `number_died`: The number of focal individuals that did not survive to adulthood.

```{r}
larvae %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "1000px", width = "800px")
```
<br></br>

# Raw data from Experiment 1: female fitness

  - `larvae_produced` The number of larvae produced (summed over all the females in the vial).
  - `number_of_laying_females` The number of females who produced the larvae. 

```{r}
females %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "1000px", width = "800px")
```
<br></br>

# Raw data from Experiment 1: male fitness

 - `focal_daughters`: The number of eclosing females that were sired by the focal males.
 - `focal_sons`: The number of eclosing males that were sired by the focal males.
 - `rival_daughters`: The number of eclosing females that were sired by the competing _Gla/CyO_ males.
 - `rival_sons`: The number of eclosing males that were sired by the competing _Gla/CyO_ males.
 
```{r}
males %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "1000px", width = "800px")
```
<br></br>

# Raw data from Experiment 2

```{r}
expt2 %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "1000px", width = "800px")
```
